Marlowe: implementing and analysing financial contracts on blockchain

### 2019-12-10 papers/marlowe2019/paper.tex
suggested changes by david

\begin{abstract}
Marlowe is a DSL for financial contracts running on the Cardano blockchain. We describe the implementation of Marlowe on blockchain, and the Marlowe Playground web-based development and simulation environment.

Contracts in Marlowe can be exhaustively analysed prior to running them, thus providing strong guarantees to participants in the contracts. The Marlowe system itself has been formally verified using the Isabelle theorem prover, establishing such properties as the conservation of money.

@@

Marlowe\footnote{The complete Marlowe project, including semantics and implementation, is available at \url{https://github.com/input-output-hk/marlowe}.} is a domain-specific language (DSL) for implementing financial contracts on blockchain: our initial target is Cardano, but it can be implemented on many DLT platforms, including Ethereum. Marlowe is embedded in Haskell, allowing users selectively to use aspects of Haskell -- typically definitions of constants and simple functions -- to express contracts more readably and succinctly. Marlowe was introduced in the paper~\cite{isola-marlowe}: Section \ref{sec:overview} below gives an overview of the language, and the changes made to it since the publication of that paper.

Marlowe is specified by a reference semantics for the language written in Haskell, and we can use that  in a number of ways. Obviously, we can interpret Marlowe contracts in Haskell itself, but we can also use that implementation, compiled into Plutus~\cite{PlutusPlatform}, to interpret  Marlowe directly on the Cardano blockchain, as explained in Section \ref{sec:implementation}. Moreover, we can execute that semantics -- translated into PureScript -- directly in a browser, to give an interactive simulation environment, the Meadow Playground, described in Section \ref{sec:playground}.

Because Marlowe is a DSL, we are able to build special purpose tools and techniques to support it. Crucially in a financial environment, we are able to \emph{exhaustively analyse contracts} without executing them, so that we can, for instance, check whether any particular contract is able to make all the payments it should: in the case it is not, we get an explicit example of how it can fail. This analysis, reported in Section \ref{sec:static}, is built into the Marlowe Playground. We are also able to use formal verification to prove properties of the implementation of Marlowe, including a guarantee that ``money in $=$ money out'' for all contracts; this is covered in Section~\ref{sec:verification}. 
% PLS: Should we divide the analysis/formalisation section in two? SJT: done

@@

Since the first publication we have revised the language design: this section gives a brief overview of the current (3.0) version of the language and its semantics.

@@

Marlowe is designed to support the execution of financial contracts on blockchain, and specifically to work on Cardano. Contracts are built by putting together a small number of constructs that in combination can be used to describe many different kinds of financial contract.

@@

\subsubsection*{Marlowe step by step}
Marlowe has five ways of building contracts. Four of these -- \texttt{Pay}, \texttt{Let}, \texttt{If} and \texttt{When} -- build a complex contract from simpler contracts, and the fifth, \texttt{Close}, is a simple contract. At each step of execution, as well as returning a new state and continuation contract, it is possible that effects, e.g.\ payments, and warnings can be generated too. We also explain Marlowe values, observations and actions, which are used to supply external information and inputs to a running contract to control how it will evolve.

\paragraph{Pay}
A payment contract \texttt{Pay a p v cont} will make a payment of value \texttt{v} from the account \texttt{a} to a payee \texttt{p}, which will be one of the contract participants or another account in the contract. Warnings will be generated if the value \texttt{v} is not positive, or if there is not enough in the account to make the payment in full. In the first case, nothing will be transferred; in the later case, a partial payment (of all the money available) is made. The contract will continue as specified by \texttt{cont}.

\paragraph{Close}
A contract \texttt{Close} provides for the contract to be closed (or terminated). The only action that is performed is to refund the contents of each account to their respective owners. This is performed one account per step, but all accounts will be refunded in a single transaction.

@@

\paragraph{Values, observations and actions}
\emph{Values} include some quantities that change with time, including the current slot number, the current balance of an account (in Lovelace), and any choices that have already been made. % ; we call these volatile values. - PLS: this definition is no longer used
Values can also be combined using addition, subtraction and negation.

\emph{Observations} are Boolean values derived by comparing values, and can be combined using the standard Boolean operators. It is also possible to observe whether any choice has been made (for a particular identified choice).

Observations will have a value at every step of execution. On the other hand, \emph{actions} happen at particular points during execution and can be
(i) depositing money, (ii)
making a choice between various alternatives, or (iii)
notifying an external or oracle value, such as the current price of a particular commodity.

\paragraph{If}
The conditional \texttt{If obs cont1 cont2} will continue as \texttt{cont1} or \texttt{cont2}, depending on the Boolean value of the observation \texttt{obs} on execution.

\paragraph{When}
This is the most complex constructor for contracts, with the form \texttt{When cases timeout cont}. It is a contract that is triggered on actions, which may or may not happen at any particular slot: the permitted actions and their consequences are described by \texttt{cases}.

@@

\paragraph{Let}
A let contract \texttt{Let id val cont} causes the expression \texttt{val} to be evaluated, and stored with the name \texttt{id}. The contract then continues as \texttt{cont}.

@@

In order to spend an existing transaction output protected by a \emph{Validator},
one must create a transaction (a \emph{PendingTx}) that has an \emph{input} that references the transaction output,
and contains a \emph{Redeemer}, such that

\begin{verbatim}
    Validator(Data Script, Redeemer, PendingTx)
\end{verbatim}
evaluates to $True$.
In order to spend a transaction output protected by a \emph{public key}, one must provide a valid signature
made with the private key corresponding to the transaction public key.

@@

The Marlowe validator script checks that a transaction contains a valid continuation
output in case a contract is not closed, by checking the hash of the output validator and state.
The continuation must have the same validator, hence the same hash of the validator script.
It should contain both the expected continuation contract and then state, evaluated by the \emph{computeTransaction}
function.

@@

When a contract evaluates to \emph{Close}, and all remaining balances must payed out
to respective owners, removing the contract from the set of unspent transaction outputs.

@@

Conclusions (draft)
\begin{itemize}
\item
Using a DSL means that we can do much better with analysis and proof than a general purpose language: so much that we can incorporate the analysis in the development environment.
\item
Defining the language by means of an executable reference semantics means that we can, as well as directly executing this semantics, generate an on-chain interpreter for it and simulate it in browser using the Haskell-like languages Plutus and PrueScript. This is made particularly straightforward by working in the subset of Haskell that directly corresponds to what translates directly into these languages.
\item
Different development modes: visual and Haskell-aided.
\item
Something about ACTUS?
\item \ldots
\end{itemize}

\medskip\noindent
Future Work (draft)
\begin{itemize}
\item Tokenising and securitising contracts.
\item Other implementations e.g. Ethereum
\item Multi-currency
\item Launching on Cardano blockchain itself 
\item \ldots
\end{itemize}

>>>>

\begin{abstract}
Marlowe is a DSL for financial contracts. We describe the implementation of Marlowe on the Cardano blockchain, and the Marlowe Playground web-based development and simulation environment.

Contracts in Marlowe can be exhaustively analysed prior to running them, thus providing strong guarantees to participants in the contract. The Marlowe system itself has been formally verified using the Isabelle theorem prover, establishing such properties as the conservation of money.

@@

Marlowe\footnote{The complete Marlowe project, including semantics and implementation, is available at \url{https://github.com/input-output-hk/marlowe}.} is a domain-specific language (DSL) for implementing financial contracts on blockchain: our initial target is Cardano, but it could be implemented on many DLT platforms, including Ethereum. Marlowe is embedded in Haskell, allowing users selectively to use aspects of Haskell -- typically definitions of constants and simple functions -- to express contracts more readably and succinctly. Section \ref{sec:overview} gives an overview of the language, and the changes made to it since it was originally introduced in ~\cite{isola-marlowe}.

Marlowe is specified by a reference semantics for the language written in Haskell, and we can use that in a number of ways. We can interpret Marlowe contracts in Haskell itself, but we can also use that implementation, compiled into Plutus~\cite{PlutusPlatform}, to interpret Marlowe directly on the Cardano blockchain, see Section \ref{sec:implementation}. We can also execute the semantics -- translated into PureScript -- directly in a browser, to give an interactive simulation environment, see Section \ref{sec:playground}.

Because Marlowe is a DSL, we are able to build special purpose tools and techniques to support it. Crucially in a financial environment, we are able to \emph{exhaustively analyse contracts} without executing them, so that we can, for instance, check whether any particular contract is able to make all the payments it should: in the case it is not, we get an explicit example of how it can fail. This analysis, reported in Section \ref{sec:static}, is built into the Marlowe Playground. Finally, we are able to use formal verification to prove properties of the implementation of Marlowe, including a guarantee that ``money in $=$ money out'' for all contracts; see Section~\ref{sec:verification}.

@@

Since the first publication, we have revised the language design: this section gives a brief overview of the current (3.0) version of the language and its semantics.

@@

Contracts are built by putting together a small number of constructs that in combination can be used to describe many different kinds of financial contract.

@@

\subsubsection*{Marlowe step by step}

Marlowe has five ways of building contracts, we call these \emph{contract constructs}. \emph{Contract constructs}, in turn, can also contain \emph{values}, \emph{observations} and \emph{actions}.

\paragraph{Values, observations and actions} are used to supply external information and inputs to a running contract to control how it will evolve.

\texttt{Value}s include some quantities that change with time, including the current slot number, the current balance of an account (in Lovelace), and any choices that have already been made. % ; we call these volatile values. - PLS: this definition is no longer used
Values can also be combined using addition, subtraction and negation.

\texttt{Observation}s are Boolean expressions that compare values, and can be combined using the standard Boolean operators. It is also possible to observe whether any choice has been made (for a particular identified choice). Observations will have a value at every step of execution.

\texttt{Action}s happen at particular points during execution and can be (i) depositing money, (ii) making a choice between various alternatives, or (iii) notifying the contract that a certain observation has become true.

\paragraph{Contract constructs} are the main building block of contracts, and there are five of them: four of these -- \texttt{Pay}, \texttt{Let}, \texttt{If} and \texttt{When} -- build a complex contract from simpler contracts, and the fifth, \texttt{Close}, is a simple contract. At each step of execution we will obtain a new state and continuation contract and, in some it is possible that effects, like payments and warnings, can be generated too.

\texttt{Pay}: A payment contract \texttt{Pay a p v cont} will make a payment of value \texttt{v} from the account \texttt{a} to a payee \texttt{p}, which will be one of the contract participants or another account in the contract. Warnings will be generated if the value \texttt{v} is not positive, or if there is not enough in the account to make the payment in full. In the first case, nothing will be transferred; in the later case, a partial payment (of all the money available) is made. The contract will continue as specified by \texttt{cont}.

\texttt{Close}: A contract \texttt{Close} provides for the contract to be closed (or terminated). The only action that is performed is to refund the contents of each account to their respective owners. This is performed one account per step, but all accounts will be refunded in a single transaction. All contracts eventually reduce to \texttt{Close}.

@@

\texttt{If}: The conditional \texttt{If obs cont1 cont2} will continue as \texttt{cont1} or \texttt{cont2}, depending on the Boolean value of the observation \texttt{obs} on execution.

\texttt{When}: This is the most complex constructor for contracts, with the form \texttt{When cases timeout cont}. It is a contract that is triggered on actions, which may or may not happen at any particular slot: the permitted actions and their consequences are described by \texttt{cases}.

@@

\texttt{Let}: A let contract \texttt{Let id val cont} causes the expression \texttt{val} to be evaluated, and stored with the name \texttt{id}. The contract then continues as \texttt{cont}.

@@

In order to spend an existing transaction output protected by a \emph{Validator},
one must create a transaction (a \emph{PendingTx}) that has an \emph{input} that references the transaction output,
and contains a \emph{Redeemer}, such that \texttt{Validator(Data Script, Redeemer, PendingTx)} evaluates to $True$.
In order to spend a transaction output protected by a \emph{public key}, one must provide a valid signature
made with the private key corresponding to the transaction public key.

@@

The Marlowe validator script checks that a transaction contains a valid continuation
output in case a contract is not closed, by checking the hash of the output validator and state.
The continuation must have the same validator, hence the same hash of the validator script.
It should contain both the expected continuation contract and the state, evaluated by the \emph{computeTransaction}
function.

@@

When a contract evaluates to \emph{Close}, all remaining balances must payed out
to respective owners, removing the contract from the set of unspent transaction outputs.

@@

Rather than aiming to be general-purpose, Marlowe is a DSL designed to support financial contracts on blockchain. We leverage its specificity in our work on static analysis and verification, where we are able to deliver much greater impact and focus than we could for a general-purpose language. We are able to shape the development and simulation environment to give stronger user support too. Moreover, Marlowe  presents a model for how other DSLs can be built in this space, supporting different domains such as provenance in the supply chain.

Defining the language by means of an executable reference semantics means that we can, as well as directly executing this semantics, generate an on-chain interpreter for it and simulate it in browser using the Haskell-like languages Plutus and PureScript. This is made particularly straightforward by working in the subset of Haskell that directly corresponds to what translates directly into these languages.

Our medium term aim is launching on Cardano blockchain itself, by which time we expect to have added multiple currencies to Marlowe, as well as making (roles in) Marlowe contracts tradeable, through tokenising contract roles.
